---
title: "Fig4C,4D"
output: html_document
date: "2026-01-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(ggplot2)
```

Run both the interneurons and motor neurons rmd and the regeneration_subclusters rmd to format, unless you have the rds files
```{r}
neural <- LoadSeuratRds(file = "neural_final.rds")
INs <- LoadSeuratRds(file = "INs.rds")
MNs <- LoadSeuratRds(file = "MNs.rds")

FeaturePlot(neural, features= c("ret"))
```

Putting in the cell type information I've learned from clustering the interneurons and motor neurons separately
```{r}
INs@meta.data$final_names <- INs@meta.data$INs_names
MNs@meta.data$final_names <- MNs@meta.data$MN_names
neural@meta.data$final_names <- neural@meta.data$neural_cell_types

INs_barcodes <- INs@meta.data[, c(1, 26)]
neural@meta.data
for (i in 1:nrow(neural@meta.data)) {
  if (rownames(neural@meta.data[i,]) %in% rownames(INs_barcodes)) {
    for (e in 1:nrow(INs_barcodes)) { 
      if (rownames(neural@meta.data[i,]) == rownames(INs_barcodes[e,])) {
        neural@meta.data[i,]$final_names <- INs_barcodes[e,]$final_names
        print("yay")
      }
      else {
        print("nay")
      }
      }
    }
  else {
    print("boo")
    }
  }

MN_barcodes <- MNs@meta.data[, c(1, 23)]
options(future.globals.maxSize = 1024^4)

for (i in 1:nrow(neural@meta.data)) {
  if (rownames(neural@meta.data[i,]) %in% rownames(MN_barcodes)) {
    for (e in 1:nrow(MN_barcodes)) { 
      if (rownames(neural@meta.data[i,]) == rownames(MN_barcodes[e,])) {
        neural@meta.data[i,]$final_names <- MN_barcodes[e,]$final_names
        print("yay")
      }
      else {
        print("nay")
      }
      }
    }
  else {
    print("boo")
    }
  }

unique(neural@meta.data$final_names)

# make whole names off of the final names- this is the entire cluster including regeneration specific ones
neural@meta.data$whole_names <- neural@meta.data$final_names
neural@meta.data$whole_names <- dplyr::recode(neural@meta.data$final_names,
                                "Regenerating prdm14- Motor Neurons" = "prdm14- Motor Neurons",
                                "Regenerating prdm14+ Motor Neurons" = "prdm14+ Motor Neurons",
                                "Regenerating V0 Glutamatergic Interneurons" = "V0 Glutamatergic Interneurons",
                                "Regenerating V2a Glutamatergic/Cholinergic Interneurons" = "V2a Glutamatergic/Cholinergic Interneurons",
                                "Regenerating Rohon Beard Neurons" = "Rohon Beard Neurons")
```

Final housekeeping
```{r}
# I don't like that there are some random MN cells in other clusters- let's pick them out
barcodes <- xSelectCells::xSelectCells(neural)
barcodes[[length(barcodes)+1]] <- xSelectCells::xSelectCells(neural)
neural2 <- subset(neural, !(neural@meta.data$orig.ident %in% barcodes))

# when happy with the selection, rename!
neural <- neural2
```

Make figure, double checking the whole name definitions
```{r}
palette <- c("Neural Progenitor Cells" = "#D6D6D6",
             "Differentiating Motor Neurons" = "#D6D6D6",
             "prdm14- Motor Neurons" = "magenta2",
             "prdm14+ Motor Neurons" = "#9D129F",
             "Roof Plate" = "#D6D6D6",
             "Differentiating Interneurons" = "#D6D6D6",
             "Floor Plate" = "#D6D6D6",
             "Rohon Beard Neurons" = "#E5AA70",
             "KA'' GABAergic Interneurons" = "steelblue1",
                "V1 GABAergic/Glycinergic Interneurons" = "mediumturquoise",
                "V0 Glutamatergic Interneurons" = "coral1",
                "dI6 Glycinergic Interneurons" = "goldenrod1",
                "V2a Glutamatergic/Cholinergic Interneurons" = "palevioletred1",
                "V2b GABAergic/Glycinergic Interneurons" = "olivedrab3",
                "KA' GABAergic/Dopaminergic Interneurons" = "dodgerblue4",
                "Intermediate KA Interneurons" = "dodgerblue3",
                "unknown Glycinergic Interneurons" = "brown")

# make a umap
DimPlot(neural, group.by = "final_names")
DimPlot(neural, group.by = "whole_names", label = FALSE, label.size = 3, repel = TRUE, cols = palette) +   theme(axis.ticks=element_blank(), axis.text=element_blank(), plot.title=element_blank(), axis.line = element_blank(), axis.title = element_blank()) +
  scale_y_continuous(limits = c(-11, 7)) + scale_x_continuous(limits = c(-9, 10)) +
  NoLegend() +
  annotate("segment", 
            x = -9, xend = -9 + c(3, 0),
            y = -11, yend = -11 + c(0, 4),
           arrow = arrow(type = "closed", length = unit(10, 'pt')))

neural[[]]
# clean up and then save
neural@meta.data <- neural@meta.data[, -which(colnames(neural@meta.data) %in% 'info')]
neural@meta.data <- neural@meta.data[, -which(colnames(neural@meta.data) %in% 'cells_of_interest')]

#saveRDS(neural, file = "neural_final.rds")
```

# Pie chart for neuron types looking in neuronal proportion for Figure 4E!
```{r}
Neuronpalette <- c("Neural Progenitor Cells" = "#D6D6D6",
             "Differentiating Motor Neurons" = "#D6D6D6",
             "prdm14- Motor Neurons" = "magenta2",
             "prdm14+ Motor Neurons" = "#9D129F",
             "Roof Plate" = "#D6D6D6",
             "Differentiating Interneurons" = "#D6D6D6",
             "Floor Plate" = "#D6D6D6",
             "Rohon Beard Neurons" = "#E5AA70",
             "KA'' GABAergic Interneurons" = "steelblue1",
              "V1 GABAergic/Glycinergic Interneurons" = "mediumturquoise",
              "V0 Glutamatergic Interneurons" = "coral1",
              "dI6 Glycinergic Interneurons" = "goldenrod1",
              "V2a Glutamatergic/Cholinergic Interneurons" = "palevioletred1",
              "V2b GABAergic/Glycinergic Interneurons" = "olivedrab3",
              "KA' GABAergic/Dopaminergic Interneurons" = "dodgerblue4",
              "Intermediate KA Interneurons" = "dodgerblue3",
              "unknown Glycinergic Interneurons" = "brown")

Neurons <- subset(x = neural, subset = neural@meta.data$actual_cell_types == "Neurons")
Neurons <- subset(x = Neurons, subset = Neurons@meta.data$whole_names != "Differentiating Interneurons")
Neurons <- subset(x = Neurons, subset = Neurons@meta.data$whole_names != "Differentiating Motor Neurons")

unique(Neurons@meta.data$whole_names)
Neurons@meta.data$whole_names <- factor(Neurons@meta.data$whole_names, levels = c(
  "unknown Glycinergic Interneurons",
  "prdm14- Motor Neurons",
  "prdm14+ Motor Neurons",
  "KA'' GABAergic Interneurons",
  "Intermediate KA Interneurons",
  "KA' GABAergic/Dopaminergic Interneurons",
  "V2b GABAergic/Glycinergic Interneurons",
  "V2a Glutamatergic/Cholinergic Interneurons",
  "V1 GABAergic/Glycinergic Interneurons",
  "V0 Glutamatergic Interneurons",
  "dI6 Glycinergic Interneurons",
  "Rohon Beard Neurons"))

# Pie chart
ggplot(Neurons@meta.data, aes(x = "", fill = whole_names)) + 
  geom_bar(position = "fill") + 
  coord_polar("y", start=0) + 
  scale_fill_manual(values=Neuronpalette) +
  facet_wrap(~Sample, ncol= 8) +
  theme_classic() +
  theme(axis.ticks=element_blank(), 
        axis.text=element_blank(), 
        plot.title=element_blank(), 
        axis.title = element_blank(), 
        strip.background = element_rect(fill = "white", color = "white"), 
        panel.border = element_blank(), 
        axis.line = element_blank()) +
  NoLegend()

# new stacked bar chart
ggplot(Neurons@meta.data, aes(x = "", fill = whole_names)) + 
  geom_bar(position = "fill") + 
  scale_fill_manual(values=Neuronpalette) +
  facet_wrap(~Sample, ncol= 8) +
  theme_classic() +
  theme(axis.ticks=element_blank(), 
        axis.text=element_blank(), 
        plot.title=element_blank(), 
        axis.title = element_blank(), 
        strip.background = element_rect(fill = "white", color = "white"), 
        panel.border = element_blank(), 
        axis.line = element_blank()) +
  NoLegend()

ggsave("4E.png", width = 7.3, height = 1.2, units = c("in"), dpi = 300)
```

Set up a dataframe to look at proportion over time - must run before succeeding chunks
```{r}
new <- neural[[]] %>% group_by(whole_names, Sample, Condition2, Age) %>% summarize(n())
neural[[]] %>% group_by(Sample) %>% summarize(n())
new$Percentage <- new$Sample
new$Percentage <- dplyr::recode(new$Percentage,
                                "un0dpa" = 781,
                                "un1dpa" = 984,
                                "un3dpa" = 702,
                                "un7dpa" = 126,
                                "0dpa" = 1115,
                                "1dpa" = 1408,
                                "3dpa" = 1726,
                                "7dpa" = 495)
new$Percentage <- new$`n()` / new$Percentage
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
new$prettyPercentage <- specify_decimal((new$Percentage*100), 2)
```

Double checking that everything holds up relative to total cell numbers - it does
```{r}
new <- neural[[]] %>% group_by(whole_names, Sample, Condition, Age) %>% summarize(n())

neural[[]] %>% group_by(Sample) %>% summarize(n())
new$Percentage <- new$Sample
new$Percentage <- dplyr::recode(new$Percentage,
                                "un0dpa" = 5771,
                                "un1dpa" = 6713,
                                "un3dpa" = 2698,
                                "un7dpa" = 476,
                                "0dpa" = 4792,
                                "1dpa" = 3858,
                                "3dpa" = 10124,
                                "7dpa" = 1420)
new$Percentage <- new$`n()` / new$Percentage
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
new$prettyPercentage <- specify_decimal((new$Percentage*100), 2)
```

By each for Figure 4C and Figure 4D!
```{r}
ggplot(new[which(new$whole_names == "prdm14- Motor Neurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "magenta2") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_prdm14-_motor_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "prdm14+ Motor Neurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "#9D129F") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_prdm14+_motor_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "KA'' GABAergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "steelblue1") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.11), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_kolmer_agduhr_interneurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

# Now, can I make it work with those missing timepoints
new
new[125, 1:4] = as.list(c("Rohon Beard Neurons", "7dpa", "Regenerating", "10dpf"))
new[125, 5] = as.integer(0)
new[125, 6] = as.double(0)
new[125, 7] = as.character(0.00)

new[126, 1:4] = as.list(c("Rohon Beard Neurons", "un7dpa", "Uninjured", "10dpf"))
new[126, 5] = as.integer(0)
new[126, 6] = as.double(0)
new[126, 7] = as.character(0.00)

ggplot(new[which(new$whole_names == "Rohon Beard Neurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "#E5AA70") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_rohon_beard_sensory_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "V1 GABAergic/Glycinergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "mediumturquoise") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_V1_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "V0 Glutamatergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "coral1") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_V0_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

new
new[127, 1:4] = as.list(c("dI6 Glycinergic Interneurons", "7dpa", "Regenerating", "10dpf"))
new[127, 5] = as.integer(0)
new[127, 6] = as.double(0)
new[127, 7] = as.character(0.00)

new[128, 1:4] = as.list(c("dI6 Glycinergic Interneurons", "un7dpa", "Uninjured", "10dpf"))
new[128, 5] = as.integer(0)
new[128, 6] = as.double(0)
new[128, 7] = as.character(0.00)

ggplot(new[which(new$whole_names == "dI6 Glycinergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "goldenrod1") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_dI6_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "V2a Glutamatergic/Cholinergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "palevioletred1") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_V2a_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "V2b GABAergic/Glycinergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "olivedrab3") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_V2b_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "Intermediate KA Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "dodgerblue3") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_intKA_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)

ggplot(new[which(new$whole_names == "KA' GABAergic/Dopaminergic Interneurons"), ], aes(x = Age, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "dodgerblue4") +
  geom_path(color = "black", group= "Condition2") +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(limits = c(0, 0.10), labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition2, ncol= 1, strip.position = "right")
ggsave("4_KA'_neurons.png", width = 1, height = 2, units = c("in"), dpi = 300)
```
