---
title: "Figure 5"
output: html_document
date: "2025-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load
```{r}
library(rstatix)
library(dplyr)
library(car)
library(GO.db)
library(Seurat)
library(readr)
library(stringr)
library(clusterProfiler)
library(scCustomize)
library(ggplot2)


neural <- LoadSeuratRds(file = "neural_final.rds")
```

My designations have changed since I original called the seurat cell types, so let's make anew column, "actual cell types" - Don't need to do if you have the final rds file
```{r}
neural@meta.data$actual_cell_types <- as.character(neural@meta.data$neural_cell_types)
neural@meta.data$actual_cell_types =
  dplyr::recode(neural@meta.data$actual_cell_types,
             "Neural Progenitor Cells" = "Neural Progenitor Cells",
             "Differentiating Motor Neurons" = "Differentiating Neurons",
             "Motor Neurons" = "Neurons",
             "Glutamatergic Interneurons" = "Neurons", 
             "prdm14+ Motor Neurons" = "Neurons",
             "lep+ Motor Neurons" = "Neurons",
             "GABA/Glycinergic Interneurons" = "Neurons",
             "Kolmer-Agduhr Interneurons" = "Neurons",
             "Roof Plate" = "Neural Progenitor Cells",
             "Differentiating Interneurons" = "Differentiating Neurons",
             "Floor Plate" = "Neural Progenitor Cells",
             "vsx2+ Interneurons" = "Neurons",
             "Rohon Beard Neurons" = "Neurons")
unique(neural@meta.data$actual_cell_types)

neural@meta.data$actual_cell_types <- factor(neural@meta.data$actual_cell_types, levels = c("Neural Progenitor Cells", "Differentiating Neurons", "Neurons"))

#saveRDS(neural, file = "neural.rds")
```

Figure S5A Phase facet UMAP
```{r}
palette <- c("#842682", "#C63D73", "#FC8A61")

DimPlot_scCustom(neural, group.by = "Phase", 
                 label = FALSE, 
                 split.by = "Sample", 
                 num_columns = 8,
                 colors_use = palette) & 
  theme(axis.ticks=element_blank(), 
        axis.text=element_blank(), 
        plot.title=element_blank(), 
        axis.line = element_blank(), 
        axis.title = element_blank()) &
 scale_y_continuous(limits = c(-11, 7)) &
 scale_x_continuous(limits = c(-9, 10)) &
 # NoLegend() &
  annotate("segment", 
            x = -9, xend = -9 + c(3, 0),
            y = -11, yend = -11 + c(0, 4),
           arrow = arrow(type = "closed", length = unit(6, 'pt')))

#ggsave("S4A.png", width = 15, height = 2, units = c("in"), dpi = 300)
```

Making some bar charts of phases for Figure 5A
```{r}
# getting the numbers
bigNeuralDataframe <- neural[[]] %>% group_by(actual_cell_types, Sample, Condition2, Age) %>% summarize(n())
neural[[]] %>% group_by(Sample) %>% summarize(n())
bigNeuralDataframe$Percentage <- bigNeuralDataframe$Sample
bigNeuralDataframe$Percentage <- dplyr::recode(bigNeuralDataframe$Percentage,
                                "un0dpa" = 781,
                                "un1dpa" = 984,
                                "un3dpa" = 702,
                                "un7dpa" = 126,
                                "0dpa" = 1115,
                                "1dpa" = 1408,
                                "3dpa" = 1726,
                                "7dpa" = 495)

bigNeuralDataframe$Percentage <- bigNeuralDataframe$`n()` / bigNeuralDataframe$Percentage
bigNeuralDataframe

Bigpalette = c("G1" = "grey20", "G0" = "grey45", "S" = "#D6D6D6")

neural@meta.data$Phase <- factor(neural@meta.data$Phase, levels = c("S", "G2M", "G1"))

# actual plot
ggplot(neural@meta.data, aes(x = Age, fill = Phase)) + 
  geom_bar(position = "fill", color = "black") + 
  facet_wrap(~Condition2) +
  scale_fill_manual(values=Bigpalette) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"), 
    strip.background = element_blank(),
    strip.text = element_blank(), panel.border = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 100, accuracy=1))

#ggsave("5A.png", width = 2.25, height = 1.9, units = c("in"), dpi = 300)
```

Plots for Figure 5B
```{r}
# back to normal figure
bigNeuralDataframe <- neural[[]] %>% group_by(actual_cell_types, Sample, Condition2, Age) %>% summarize(n())
bigNeuralDataframe$Score <- bigNeuralDataframe$Sample
bigNeuralDataframe$Score <- dplyr::recode(bigNeuralDataframe$Score,
  "un0dpa" = "0.0196694904",			
  "un1dpa" =	"0.0008685252",			
  "un3dpa" =	"-0.0094266682",			
  "un7dpa" = "-0.0116251256",			
  "0dpa" =	"0.0032123799",		
  "1dpa" =	"-0.0144049534",			
  "3dpa" =	"0.0329315755",			
  "7dpa" =	"-0.0087068619")

bigNeuralDataframe$Score3 <- bigNeuralDataframe$Sample
bigNeuralDataframe$Score3 <- dplyr::recode(bigNeuralDataframe$Score3,
  "un0dpa" = "-3.727419e-05",			
  "un1dpa" =	"-2.745690e-02",			
  "un3dpa" =	"-2.684760e-02",			
  "un7dpa" = "-3.190311e-02",			
  "0dpa" =	"1.276520e-02",		
  "1dpa" =	"-2.951595e-02",			
  "3dpa" =	"-2.637207e-02",			
  "7dpa" =	"1.820343e-02")
                                               
neural[[]] %>% group_by(Sample) %>% summarize(n())
bigNeuralDataframe$Percentage <- bigNeuralDataframe$Sample
bigNeuralDataframe$Percentage <- dplyr::recode(bigNeuralDataframe$Percentage,
                                "un0dpa" = 781,
                                "un1dpa" = 984,
                                "un3dpa" = 702,
                                "un7dpa" = 126,
                                "0dpa" = 1115,
                                "1dpa" = 1408,
                                "3dpa" = 1726,
                                "7dpa" = 495)

bigNeuralDataframe$Percentage <- bigNeuralDataframe$`n()` / bigNeuralDataframe$Percentage

Bigpalette = c("Neural Progenitor Cells" = "#D6D6D6", "Differentiating Neurons" = "grey45", "Neurons" = "grey20")

# actual plot
ggplot(neural@meta.data, aes(x = Age, fill = actual_cell_types)) + 
  geom_bar(position = "fill", color = "black") + 
  facet_wrap(~Condition2) +
  scale_fill_manual(values=Bigpalette) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    axis.text.x = element_blank(), 
    panel.background = element_rect(fill = "white"), 
    strip.background = element_blank(),
    strip.text = element_blank(), panel.border = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 100, accuracy=1))

#ggsave("5B.png", width = 2.25, height = 1.9, units = c("in"), dpi = 300)
```

SET UP GO TERMS
```{r}
f <- function(x, pos) dplyr::select(x, X3, X5)
go_ids <- read_tsv_chunked("QuickGO-annotations-1750272346286-20250618.gaf", callback = DataFrameCallback$new(f), col_names = FALSE)
colnames(go_ids) <- c("gene", "GOID")

df <- AnnotationDbi::select(GO.db, columns=c("GOID","TERM", "ONTOLOGY"), keys = c("BP", "CC", "MF"), keytype="ONTOLOGY")
go_ids <- left_join(go_ids, df, join_by("GOID"))
go_ids <- go_ids %>% filter(!is.na(TERM))
go_ids <- unique.data.frame(go_ids)
go_ids

#Creates table of all differentially expressed genes between matched timepoints in regenerating and uninjured #conditions in three major cell states (NPC, Differentiating Neurons, Neurons)

neural$sample.state <- paste(neural$Sample, neural$actual_cell_types, sep = "_")
neural$sample.state <- factor(neural$sample.state)
Idents(neural) <- "sample.state"
sample.states <- levels(neural$sample.state)
de_genes <- data.frame(gene = c(), log2FC = c())
index <- 1

for (i in 1:12) {
  neural.de <- FindMarkers(object = neural,
                         ident.1 = sample.states[index],
                         ident.2 = sample.states[index+12], 
                         verbose = FALSE, min.pct = 0.3, 
                         min.cells.group = 0, assay = "RNA")
  genes <- data.frame(rownames(neural.de), neural.de$avg_log2FC)
  colnames(genes) <- c("gene", "log2FC")
  genes$direction <- "upregulated"
  genes$direction[genes$log2FC < 0] <- "downregulated"
  genes$Age <- paste0(substring(sample.states[index], 1, 1), "dpa")
  genes$cell_type <- substring(sample.states[index], 6, nchar(sample.states[index]))
  de_genes <- rbind(de_genes, genes)
  index <- index + 1
}


#Performs GO analysis on each sample and identifies upregulated and downregulated terms enriched at each #timepoint amongst differentially expressed genes with a fold change >= 1.5 (NPC, Differentiating Neurons, #Neurons)

enrich_genes <- de_genes %>% filter(abs(log2FC) >= log2(1.5))
formula_res <- compareCluster(gene~direction+Age+cell_type, data=enrich_genes, fun="enricher", TERM2GENE=go_ids[go_ids$ONTOLOGY == "BP", c("TERM", "gene")])
temp <- data.frame(paste(formula_res@compareClusterResult[["Description"]],formula_res@compareClusterResult[["Cluster"]]), formula_res@compareClusterResult[["geneID"]])
colnames(temp) <- c("Description", "geneID")
temp$geneID <- strsplit(temp$geneID, "/")
temp1 <- as.list(temp$geneID)
names(temp1) <- temp$Description
temp1 <- stack(temp1)
temp1$Age <- str_extract(temp1$ind, ".dpa")
temp1
temp1$cell_type <- str_extract(temp1$ind, "Neurons|Differentiating Neurons|Neural Progenitor Cells")
temp1 <- left_join(temp1, de_genes %>% dplyr::select(gene, log2FC, Age, cell_type), by = join_by(values == gene, Age == Age, cell_type == cell_type))
temp <- temp1 %>% dplyr::group_by(ind) %>% summarize(avglog2FC = mean(log2FC))
enriched_go <- data.frame(formula_res)
enriched_go <- mutate(enriched_go, avglog2foldChange = temp$avglog2FC)
go_id_map <- data.frame(go_ids$GOID, go_ids$TERM)
go_id_map <- unique.data.frame(go_id_map)
enriched_go$ID <- plyr::mapvalues(enriched_go$ID,go_id_map$go_ids.TERM,go_id_map$go_ids.GOID,warn_missing = F)
```

INDIVIDUAL GO TERMS (cell division, Figure 5C)
```{r}
genes <- Features(neural)

genes_of_interest <- go_ids %>% 
  filter(TERM %in% c("cell division")) %>% 
  dplyr::select(gene)
genes_of_interest <- unique.data.frame(genes_of_interest)
genes_of_interest <- list(intersect(genes_of_interest$gene,genes))
neural <- AddModuleScore(neural, features = genes_of_interest, name = "interest")

# visualizing score on the UMAP
FeaturePlot_scCustom(neural, reduction = "umap", 
            pt.size = 1, 
            features = "interest1", 
            order = TRUE,
            split.by = "Condition2",
            num_columns = 2) & 
  scale_color_distiller(palette = "RdYlBu")

# cell division GO term over actual_cell_types and Sample
ggplot(neural@meta.data, aes(x = actual_cell_types, y = interest1, fill = Sample)) + 
  geom_boxplot(alpha = 0.1) + 
  theme(panel.background = element_blank(), axis.line = element_line(color = "black"), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# Figure 5C
Condpalette = c("Uninjured" = "white", "Regenerating" = "grey")

ggplot(neural@meta.data, aes(x = Sample, y = interest1, fill = Condition2, lwd = "black")) + 
  scale_fill_manual(values = Condpalette) +
  geom_boxplot(lwd = 0.4, outlier.shape=NA) +
  ylab("Cell Division") +
  theme(panel.background = element_blank(), 
        axis.line = element_line(color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) +
  NoLegend()

#ggsave("5C.png", width = 2, height = 2.5, units = c("in"), dpi = 300)
```

# Stats for celldivision (interest1)
```{r}
meta <- neural[[]] %>% dplyr::select(Sample, Condition2, interest1)
meta <- meta[which(meta$Condition2 == "Regenerating"),]
#Statistics for module plot
atab <- aov(interest1~Sample, data = meta)
#Levene test for equal variances
leveneTest(interest1~Sample,  meta)
# Extract the residuals
aov_residuals <- residuals(object = atab)
# Run Shapiro-Wilk test for normality
shapiro.test(x = aov_residuals)
# Kruskal-Wallis test = non-parametric multiple comparisons
#kruskal.test(interest1~Condition, data = meta)
#Wilcox test = non-parametric pairwise comparisons

pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "1dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "3dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "7dpa")))
```

```{r}
meta <- neural[[]] %>% dplyr::select(Sample, Condition2, interest1)
#meta <- meta[which(meta$Condition2 == "Uninjured"),]
#Statistics for neuron_differentiation module plot
atab <- aov(interest1~Sample, data = meta)
#Levene test for equal variances
leveneTest(interest1~Sample,  meta)
# Extract the residuals
aov_residuals <- residuals(object = atab)
# Run Shapiro-Wilk test for normality
shapiro.test(x = aov_residuals)
# Kruskal-Wallis test = non-parametric multiple comparisons
#kruskal.test(interest1~Condition, data = meta)
#Wilcox test = non-parametric pairwise comparisons

pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "1dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "3dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "7dpa")))
```

# Figure S5B - cell division UMAPS over Samples
```{r}
FeaturePlot_scCustom(neural, features = "interest1", 
                 label = FALSE, 
                 split.by = "Sample", 
                 num_columns = 8,
                 colors_use = viridis_magma_dark_high) & 
  theme(axis.ticks=element_blank(), 
        axis.text=element_blank(), 
        plot.title=element_blank(), 
        axis.line = element_blank(), 
        axis.title = element_blank()) &
 scale_y_continuous(limits = c(-11, 7)) &
 scale_x_continuous(limits = c(-9, 10)) &
 # NoLegend() &
  annotate("segment", 
            x = -9, xend = -9 + c(3, 0),
            y = -11, yend = -11 + c(0, 4),
           arrow = arrow(type = "closed", length = unit(6, 'pt')))

#ggsave("S5B.png", width = 15, height = 2, units = c("in"), dpi = 300)
```

Figure S5E
```{r}
neu_de_genes <- de_genes %>% filter(direction == "upregulated", log2FC > log2(1.5)) %>% dplyr::select(gene)

pseudo_neurons <- AverageExpression(subset(neural), 
                                             assays = "RNA", 
                                             return.seurat = T, 
                                             group.by = c("Sample"))

agg_neurons <- pseudo_neurons[["RNA"]]$scale.data

neu_diff_genes <- go_ids %>% 
  filter(TERM %in% c("cell division")) %>% 
  dplyr::select(gene)

genes_of_interest <- list(intersect(neu_diff_genes[[1]],neu_de_genes$gene))

agg_neurons[rownames(agg_neurons) %in% genes_of_interest[[1]],]

palette <- colorRampPalette(c("lightgoldenrod1", "red3"))
heatmap(agg_neurons[rownames(agg_neurons) %in% genes_of_interest[[1]],],
          Colv = NA,
          col = palette(100),
          keep.dendro = FALSE)
```

INDIVIDUAL GO TERMS (neuron differentiation, Figure 5D)
```{r}
genes <- Features(neural)

genes_of_interest <- go_ids %>% 
  filter(TERM %in% c("neuron differentiation")) %>% 
  dplyr::select(gene)
genes_of_interest <- unique.data.frame(genes_of_interest)
genes_of_interest <- list(intersect(genes_of_interest$gene,genes))
neural <- AddModuleScore(neural, features = genes_of_interest, name = "interest")

# visualizing score on the UMAP
FeaturePlot_scCustom(neural, reduction = "umap", 
            pt.size = 1, 
            features = "interest1", 
            order = TRUE,
            split.by = "Condition2",
            num_columns = 2) & 
  scale_color_distiller(palette = "RdYlBu")

# cell division GO term over actual_cell_types and Sample
ggplot(neural@meta.data, aes(x = actual_cell_types, y = interest1, fill = Sample)) + 
  geom_boxplot(alpha = 0.1) + 
  theme(panel.background = element_blank(), axis.line = element_line(color = "black"), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# Figure 5C
Condpalette = c("Uninjured" = "white", "Regenerating" = "grey")

ggplot(neural@meta.data, aes(x = Sample, y = interest1, fill = Condition2, lwd = "black")) + 
  scale_fill_manual(values = Condpalette) +
  geom_boxplot(lwd = 0.4, outlier.shape=NA) +
  ylab("Cell Division") +
  theme(panel.background = element_blank(), 
        axis.line = element_line(color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) +
  NoLegend()

#ggsave("5D.png", width = 2, height = 2.5, units = c("in"), dpi = 300)
```

# Stats for neurondifferentiation (interest1)
```{r}
meta <- neural[[]] %>% dplyr::select(Sample, Condition2, interest1)
#meta <- meta[which(meta$Condition2 == "Regenerating"),]
#Statistics for module plot
atab <- aov(interest1~Sample, data = meta)
#Levene test for equal variances
leveneTest(interest1~Sample,  meta)
# Extract the residuals
aov_residuals <- residuals(object = atab)
# Run Shapiro-Wilk test for normality
shapiro.test(x = aov_residuals)
# Kruskal-Wallis test = non-parametric multiple comparisons
#kruskal.test(interest1~Condition, data = meta)
#Wilcox test = non-parametric pairwise comparisons

pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "1dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "3dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("0dpa", "7dpa")))
```

```{r}
meta <- neural[[]] %>% dplyr::select(Sample, Condition2, interest1)
meta <- meta[which(meta$Condition2 == "Uninjured"),]
#Statistics for neuron_differentiation module plot
atab <- aov(interest1~Sample, data = meta)
#Levene test for equal variances
leveneTest(interest1~Sample,  meta)
# Extract the residuals
aov_residuals <- residuals(object = atab)
# Run Shapiro-Wilk test for normality
shapiro.test(x = aov_residuals)
# Kruskal-Wallis test = non-parametric multiple comparisons
#kruskal.test(interest1~Condition, data = meta)
#Wilcox test = non-parametric pairwise comparisons

pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "un1dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "un3dpa")))
pairwise_wilcox_test(meta, interest1~Sample, comparisons = list(c("un0dpa", "un7dpa")))
```

# Figure S5C - neuron differentiation UMAPS over Samples
```{r}
FeaturePlot_scCustom(neural, features = "interest1", 
                 label = FALSE, 
                 split.by = "Sample", 
                 num_columns = 8,
                 colors_use = viridis_magma_dark_high) & 
  theme(axis.ticks=element_blank(), 
        axis.text=element_blank(), 
        plot.title=element_blank(), 
        axis.line = element_blank(), 
        axis.title = element_blank()) &
 scale_y_continuous(limits = c(-11, 7)) &
 scale_x_continuous(limits = c(-9, 10)) &
 # NoLegend() &
  annotate("segment", 
            x = -9, xend = -9 + c(3, 0),
            y = -11, yend = -11 + c(0, 4),
           arrow = arrow(type = "closed", length = unit(6, 'pt')))

#ggsave("S5B.png", width = 15, height = 2, units = c("in"), dpi = 300)
```

Figure S5F
```{r}
neu_de_genes <- de_genes %>% filter(direction == "upregulated", log2FC > log2(1.5)) %>% dplyr::select(gene)

pseudo_neurons <- AverageExpression(subset(neural), 
                                             assays = "RNA", 
                                             return.seurat = T, 
                                             group.by = c("Sample"))

agg_neurons <- pseudo_neurons[["RNA"]]$scale.data

neu_diff_genes <- go_ids %>% 
  filter(TERM %in% c("neuron differentiation")) %>% 
  dplyr::select(gene)

genes_of_interest <- list(intersect(neu_diff_genes[[1]],neu_de_genes$gene))

agg_neurons[rownames(agg_neurons) %in% genes_of_interest[[1]],]

palette <- colorRampPalette(c("lightgoldenrod1", "red3"))
heatmap(agg_neurons[rownames(agg_neurons) %in% genes_of_interest[[1]],],
          Colv = NA,
          col = palette(100),
          keep.dendro = FALSE)
```

Figure S5G
```{r}
DotPlot(neural, 
        features = c("neurog1", "neurog2", "ascl1"), 
        group.by = "Sample",
        cols = c("lightgoldenrod1", "red3")) +
  coord_flip()

ggsave("S5G.png", width = 5, height = 7, units = c("in"), dpi = 300)
```