---
title: "Fig7"
output: html_document
date: "2026-01-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Commented packages are only needed when setting up the initial cellchat objects
```{r}
library(CellChat)
library(patchwork)
library(Seurat)
library(readr)
library(annotate)
library(org.Mm.eg.db)
#options(stringsAsFactors = FALSE)
#BiocManager::install("org.Mm.eg.db")
```

Load dataset
```{r}
neural <- LoadSeuratRds(file = "neural_final.rds")
```

This code chunk takes the mouse ortholog mapping data from xenbase and uses it to convert the gene names in the seurat object
```{r}
ortho <- read_tsv("XenbaseGeneMouseOrthologMapping.txt", col_names = FALSE)
data.input <- GetAssayData(neural, assay = "RNA", layer = "data")# Get normalized data
data.input <- data.input[intersect(rownames(data.input),ortho$X3),]#subset to rat genes that are 1:1 orthologs of mouse
rownames(data.input) <- plyr::mapvalues(rownames(data.input),ortho$X3,ortho$X1,warn_missing = F)# Convert rat to mouse
rownames(data.input) <- getSYMBOL(rownames(data.input), data = "org.Mm.eg")
```

This loops over all the samples and creates a cellchat object for each of them. Requires a large amount of memory, took about 1 hour for the neural subset
```{r}
unique(neural@meta.data$regenerating_names)
samples <- levels(neural$Sample)
neural$samples <- neural$Sample

for(k in 1:8) {
  data.input <- GetAssayData(neural, assay = "RNA", slot = "data")# Get normalized data
  data.input <- data.input[intersect(rownames(data.input),ortho$X3),]#subset to rat genes that are 1:1 orthologs of mouse
  rownames(data.input) <- plyr::mapvalues(rownames(data.input),ortho$X3,ortho$X1,warn_missing = F)# Convert rat to mouse
  rownames(data.input) <- getSYMBOL(rownames(data.input), data = "org.Mm.eg")
  meta <- neural[[]]
  cell.use = rownames(meta)[meta$Sample == samples[k]]

# Subset the input data for CellChat analysis
  data.input = data.input[,cell.use]

  meta = meta %>% subset(rownames(meta) %in% cell.use)

  cellchat <- createCellChat(object = data.input, meta = meta, group.by = "regenerating_names")#make cell chat object
  cellchat@idents = droplevels(cellchat@idents)
  CellChatDB <- CellChatDB.mouse
  cellchat@DB <- CellChatDB

  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
  cellchat <- identifyOverExpressedGenes(cellchat, do.fast = FALSE)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  cellchat <- computeCommunProb(cellchat, type = "triMean")
  cellchat <- filterCommunication(cellchat, min.cells = 5)
  cellchat <- computeCommunProbPathway(cellchat)
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

  saveRDS(cellchat, file = paste0("cellchat_", samples[k], "_neural_regenNames.rds"))
}
```

This code chunk takes the sample cellchat objects from the working directory and merges them into a single cellchat object containing data for all cell types collected at the 3dpa timepoint.
```{r}
# make list of files
files = list.files(pattern = "neural_regenNames.rds")
files
# as long as they contain the words neural and rds, but not merged
files <- files[files != "neural_regenNames.rds"]
files <- files[files != "cellchat_neural_regenNames.rds"]

# 3dpa has all the possible relevant cell types, let's rename it to full_dataset so we can use it later
full_dataset <- readRDS("cellchat_3dpa_neural_regenNames.rds")
# take the full dataset and collect idents for group.new
group.new = levels(full_dataset@idents)
#group.new[40] <- "dnah3+ Cells"
#group.new[41] <- "cnfn+ Epidermal Cells"

#full_dataset <- readRDS("cellchat_un0dpa_tail.rds")
#levels(full_dataset@idents)

# put files in to object list?
object.list <- lapply(files, readRDS)
# without the name .rds
name <- sub("\\.rds","",files)

# lift cellchat items so we can merge
for (i in 1:length(files)){
  rds <- object.list[[i]]
  rds <- liftCellChat(rds, group.new)
  object.list[[i]] <- rds
}
names(object.list) <- name
# merge
cellchat <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
# check
cellchat@idents
#saveRDS(cellchat, "cellchat_merged_neural_regen_names.rds")
```

Bar graph with the number of detected interactions at 1 dpa # Figure S5A
```{r}
cellchat1dpa <- readRDS("cellchat_1dpa_neural_regenNames.rds")

groupSize <- as.numeric(table(cellchat1dpa@idents))

pdf(file ="1dpa_weights.pdf", width = 5, height = 5)
netVisual_circle(cellchat1dpa@net$weight, 
                 vertex.weight = groupSize, 
                 weight.scale = T, 
                 label.edge= F, 
                 title.name = "Interaction weights/strength")
dev.off()
```

Heatmap showing the contribution of signals to cell groups in terms of outgoing or incoming signaling
# Figures S5C-F
```{r}
iNeuron_incoming <- c("NCAM", "L1CAM", "SEMA3", "EPHB", "IGF", "CNTN", "BMP", "NPR2", "KIT")
iNeuron_outgoing <- c("NRXN", "NCAM", "L1CAM", "AGRN", "NRG", "APP", "CNTN", "NOTCH", "MK")
signaling_list_incoming <- c("NRXN", "ADGRL", "Glutamate", "NCAM", "PTPR")
signaling_list_outgoing <- c("NCAM", "L1CAM", "AGRN", "EPHA", "LEP")


netAnalysis_signalingRole_heatmap(
  cellchat1dpa,
  signaling = signaling_list_incoming,
  pattern = c("incoming"),
  slot.name = "netP",
  color.use = NULL,
  color.heatmap = "BuGn",
  title = NULL,
  width = 7,
  height = 10,
  font.size = 8,
  font.size.title = 10,
  cluster.rows = FALSE,
  cluster.cols = FALSE
)

netAnalysis_signalingRole_heatmap(
  cellchat1dpa,
  signaling = iNeuron_incoming,
  pattern = c("incoming"),
  slot.name = "netP",
  color.use = NULL,
  color.heatmap = "BuGn",
  title = NULL,
  width = 7,
  height = 10,
  font.size = 8,
  font.size.title = 10,
  cluster.rows = FALSE,
  cluster.cols = FALSE
)

netAnalysis_signalingRole_heatmap(
  cellchat1dpa,
  signaling = iNeuron_outgoing,
  pattern = c("outgoing"),
  slot.name = "netP",
  color.use = NULL,
  color.heatmap = "BuGn",
  title = NULL,
  width = 7,
  height = 10,
  font.size = 8,
  font.size.title = 10,
  cluster.rows = FALSE,
  cluster.cols = FALSE
)

netAnalysis_signalingRole_heatmap(
  cellchat1dpa,
  signaling = signaling_list_outgoing,
  pattern = c("outgoing"),
  slot.name = "netP",
  color.use = NULL,
  color.heatmap = "BuGn",
  title = NULL,
  width = 7,
  height = 10,
  font.size = 8,
  font.size.title = 10,
  cluster.rows = FALSE,
  cluster.cols = FALSE
)

```

# Figure S5B
Circle diagram indicating pathway network interactions between cell types at each time point. Chord diagram can't be directly plotted in R, need to save to pdf. 
```{r}
pathways.show <- c("L1CAM")
par(mfrow = c(2,5), xpd=TRUE)
# 1 dpa
object.list[[2]]

pdf(file ="cellchatncam_1dpa.pdf", width = 50, height = 50)
par(mfrow = c(2,4), xpd=TRUE)
netVisual_aggregate(object.list[[2]], 
                      signaling = pathways.show, 
                      layout = "chord")
dev.off()
pdf(file ="cellchatncam_3dpa.pdf", width = 50, height = 50)
par(mfrow = c(2,4), xpd=TRUE)
netVisual_aggregate(object.list[[3]], 
                      signaling = pathways.show, 
                      layout = "chord")
dev.off()

pdf(file ="cellchatncam_7dpa.pdf", width = 50, height = 50)
par(mfrow = c(2,4), xpd=TRUE)
netVisual_aggregate(object.list[[4]], 
                      signaling = pathways.show, 
                      layout = "chord")
dev.off()

```