---
title: "neural_figures"
output: html_document
date: "2025-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load
```{r}
library(ggplot2)
library(dplyr)
library(Seurat)
library(scCustomize)
library(xSelectCells)
```
#devtools::install_github('vikkki/xSelectCells@main', force = TRUE)

Load dataset and level Samples
```{r}
neural <- LoadSeuratRds(file = "neural_final.rds")
INs <- LoadSeuratRds(file = "INs.rds")
MNs <- LoadSeuratRds(file = "MNs.rds")
```
Go run the INs and MNs sets and come back

Make palette
```{r}
# make a palette
palette <- c("Neural Progenitor Cells" = "#D6D6D6",
             "Differentiating Motor Neurons" = "#EBC2F3",
             "isl2+ Motor Neurons" = "magenta2",
             "Excitatory Interneurons" = "green4", 
             "prdm14+ Motor Neurons" = "#AD03AE",
             "Inhibitory Interneurons" = "green3",
             "Kolmer-Agduhr Interneurons" = "green1",
             "Roof Plate" = "#EBEBEB",
             "Differentiating Interneurons" = "#B7E9BA",
             "Floor Plate" = "#C1C1C1",
             "Rohon Beard Neurons" = "goldenrod1")
```

Reprocess MNs
```{r}
# recluster MNs to identify regenerating cells
MNs <- FindClusters(MNs, resolution = 5)

# label
MNs@meta.data$MN_names <- as.character(MNs@meta.data$seurat_clusters)
MNs@meta.data$MN_names =
  dplyr::recode(MNs@meta.data$MN_names,
                "0" = "isl2+ Motor Neurons",
                "1" = "isl2+ Motor Neurons",
                "2" = "isl2+ Motor Neurons",
                "3" = "isl2+ Motor Neurons",
                "4" = "Regenerating isl2+ Motor Neurons",
                "5" = "isl2+ Motor Neurons",
                "6" = "Differentiating Motor Neurons",
                "7" = "isl2+ Motor Neurons", 
                "8" = "isl2+ Motor Neurons",
                "9" = "isl2+ Motor Neurons",
                "10" = "isl2+ Motor Neurons",
                "11" = "isl2+ Motor Neurons",
                "12" = "Differentiating Motor Neurons",
                "13" = "Differentiating Motor Neurons",
                "14" = "prdm14+ Motor Neurons",
                "15" = "isl2+ Motor Neurons",
                "16" = "prdm14+ Motor Neurons",
                "17" = "Differentiating Motor Neurons",
                "18" = "prdm14+ Motor Neurons",
                "19" = "Regenerating isl2+ Motor Neurons",
                "20" = "Differentiating Motor Neurons",
                "21" = "Differentiating Motor Neurons",
                "22" = "isl2+ Motor Neurons",
                "23" = "isl2+ Motor Neurons",
                "24" = "Regenerating isl2+ Motor Neurons",
                "25" = "isl2+ Motor Neurons",
                "26" = "prdm14+ Motor Neurons",
                "27" = "isl2+ Motor Neurons",
                "28" = "isl2+ Motor Neurons",
                "29" = "isl2+ Motor Neurons",
                "30" = "isl2+ Motor Neurons",
                "31" = "isl2+ Motor Neurons")

DimPlot(MNs, group.by = "seurat_clusters", label = TRUE)
DimPlot(MNs, group.by = "MN_names", label = TRUE)
DimPlot(MNs, group.by = "neural_cell_types", label = TRUE)
DimPlot(MNs, group.by = "Sample", label = TRUE)

# make dataframe so we can look at percentages of these cell types
MNsDF <- MNs[[]] %>% group_by(MN_names, Sample, Condition, Age) %>% summarize(n())
MNsDF$Percentage <- MNsDF$Sample
MNsDF$Percentage <- dplyr::recode(MNsDF$Percentage,
                                "un0dpa" = 781,
                                "un1dpa" = 984,
                                "un3dpa" = 702,
                                "un7dpa" = 126,
                                "0dpa" = 1115,
                                "1dpa" = 1408,
                                "3dpa" = 1726,
                                "7dpa" = 495)
MNsDF$Percentage <- MNsDF$`n()` / MNsDF$Percentage
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
MNsDF$prettyPercentage <- specify_decimal((MNsDF$Percentage*100), 2)

# double check pattern
palette <- c("isl2+ Motor Neurons" = "magenta2", "Regenerating isl2+ Motor Neurons" = "magenta4")
MNsDF2 <- subset(MNsDF, MNsDF$MN_names == "Regenerating isl2+ Motor Neurons" | MNsDF$MN_names == "isl2+ Motor Neurons")

ggplot(MNsDF2, aes(x = Age, y = Percentage, fill = MN_names)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = palette) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition, ncol= 1, strip.position = "right")
#ggsave("isl2MNs.png", width = 1, height = 2, units = c("in"), dpi = 300)
```

Now the prdm14+ cells do not want to cluster into uninjured and regenerating cells, so I have to switch to clustering it manually
```{r}
# manually isolate regenerating lepr+ prdm14+ cells
barcodes <- xSelectCells::xSelectCells(MNs)
MNs$cells_of_interest <- ifelse(rownames(MNs@meta.data) %in% barcodes, "Regenerating prdm14+ Motor Neurons", "prdm14+ Motor Neurons")

MNsP <- subset(x = MNs, subset = (MNs@meta.data$MN_names == "prdm14+ Motor Neurons" | MNs@meta.data$MN_names == "Regenerating prdm14+ Motor Neurons"))

MNsDF3 <- MNsP[[]] %>% group_by(Sample, Condition, Age, cells_of_interest) %>% summarize(n())
MNsDF3$Percentage <- MNsDF3$Sample
MNsDF3$Percentage <- dplyr::recode(MNsDF3$Percentage,
                                "un0dpa" = 781,
                                "un1dpa" = 984,
                                "un3dpa" = 702,
                                "un7dpa" = 126,
                                "0dpa" = 1115,
                                "1dpa" = 1408,
                                "3dpa" = 1726,
                                "7dpa" = 495)
MNsDF3$Percentage <- MNsDF3$`n()` / MNsDF3$Percentage
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
MNsDF3$prettyPercentage <- specify_decimal((MNsDF3$Percentage*100), 2)

# double check pattern
palette3 <- c("prdm14+ Motor Neurons" = "magenta2", "Regenerating prdm14+ Motor Neurons" = "magenta4")

ggplot(MNsDF3, aes(x = Age, y = Percentage, fill = cells_of_interest)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = palette3) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(legend.position = "none", 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 8, color = "black"), 
    panel.background = element_rect(fill = "white", colour = "black"), 
    panel.border = element_rect(linetype = "solid", fill = "NA"),  
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 100, accuracy=1)) +
  facet_wrap(~Condition, ncol= 1, strip.position = "right")
#ggsave("prdm14MNs.png", width = 1, height = 2, units = c("in"), dpi = 300)
```

Let's add this labelling to the actual MNs object and save it
```{r}
MNs@meta.data <- MNs@meta.data %>% mutate(info=paste(MN_names, cells_of_interest, sep = "_"))
MNs@meta.data
MNs@meta.data$MN_names =
  dplyr::recode(MNs@meta.data$info,
                "isl2+ Motor Neurons_prdm14+ Motor Neurons" = "isl2+ Motor Neurons",
                "prdm14+ Motor Neurons_prdm14+ Motor Neurons" = "prdm14+ Motor Neurons",
                "Differentiating Motor Neurons_prdm14+ Motor Neurons" = "Differentiating Motor Neurons",
                "prdm14+ Motor Neurons_Regenerating prdm14+ Motor Neurons" = "Regenerating prdm14+ Motor Neurons",
                "Regenerating isl2+ Motor Neurons_Regenerating prdm14+ Motor Neurons" = "Regenerating prdm14+ Motor Neurons",
                "Regenerating isl2+ Motor Neurons_prdm14+ Motor Neurons" = "Regenerating isl2+ Motor Neurons")

MNs@meta.data$MN_names <- factor(MNs@meta.data$MN_names, levels = c(
             "prdm14+ Motor Neurons",
             "isl2+ Motor Neurons",
             "Regenerating prdm14+ Motor Neurons",
             "Regenerating isl2+ Motor Neurons",
             "Differentiating Motor Neurons"))

MNs@meta.data <- MNs@meta.data[,!names(MNs@meta.data) %in% c("cells_of_interest", "info", "sfrp4pos", "cell_types", "NoGood", "nFeat_good", "final_names")]
MNs@meta.data
#saveRDS(MNs, file = "MNs.rds")

DotPlot_scCustom(MNs, features = c("lepr"), group.by = "MN_names")
DimPlot(MNs, group.by = "final_names", label = TRUE)
```

Okay now look at the interneurons!
```{r}
DimPlot(INs, group.by = "Sample", label = FALSE)

# Yeah these are also not going to nicely cluster
DimPlot(INs, group.by = "seurat_clusters", label = FALSE)

# V2a
barcodes <- xSelectCells::xSelectCells(INs)
INs$cells_of_interest <- ifelse(rownames(INs@meta.data) %in% barcodes, "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons", "V2a/b Glutamatergic/Cholinergic Interneurons")
INs@meta.data <- INs@meta.data %>% mutate(info=paste(INs_names, cells_of_interest, sep = "_"))
INs@meta.data$INs_names =
  dplyr::recode(INs@meta.data$info,
                "KA'' GABAergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "KA'' GABAergic Interneurons",
                "V1 GABAergic/Glycinergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "V1 GABAergic/Glycinergic Interneurons",
                "V0 Glutamatergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "V0 Glutamatergic Interneurons",
                "Regenerating V0 Glutamatergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "Regenerating V0 Glutamatergic Interneurons",
                "V2a/b Glutamatergic/Cholinergic Interneurons_Regenerating V2a/b Glutamatergic/Cholinergic Interneurons" = "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons",
                "V2a/b Glutamatergic/Cholinergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "V2a/b Glutamatergic/Cholinergic Interneurons",
                "dI6 Glycinergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "dI6 Glycinergic Interneurons",
                "V2b/s GABAergic/Glycinergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "V2b/s GABAergic/Glycinergic Interneurons",
                "KA' GABAergic/Dopaminergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "KA' GABAergic/Dopaminergic Interneurons",
                "Intermediate KA Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "Intermediate KA Interneurons",
                "unknown Glycinergic Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "unknown Glycinergic Interneurons",
                "Differentiating Interneurons_V2a/b Glutamatergic/Cholinergic Interneurons" = "Differentiating Interneurons")

DimPlot(INs, group.by = "INs_names", label = FALSE)

# V0
barcodes <- xSelectCells::xSelectCells(INs2)

INs@meta.data <- INs@meta.data[,!names(INs@meta.data) %in% c("info", "cells_of_interest")]
INs$cells_of_interest <- ifelse(rownames(INs@meta.data) %in% barcodes, "Regenerating V0 Glutamatergic Interneurons", "V0 Glutamatergic Interneurons")

INs@meta.data <- INs@meta.data %>% mutate(info=paste(final_names, cells_of_interest, sep = "_"))
INs@meta.data$INs_names =
  dplyr::recode(INs@meta.data$info,
                "KA'' GABAergic Interneurons_V0 Glutamatergic Interneurons" = "KA'' GABAergic Interneurons",
                "V1 GABAergic/Glycinergic Interneurons_V0 Glutamatergic Interneurons" = "V1 GABAergic/Glycinergic Interneurons",
                "V0 Glutamatergic Interneurons_V0 Glutamatergic Interneurons" = "V0 Glutamatergic Interneurons",
                "V0 Glutamatergic Interneurons_Regenerating V0 Glutamatergic Interneurons" = "Regenerating V0 Glutamatergic Interneurons",
                "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons_V0 Glutamatergic Interneurons" = "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons",
                "dI6 Glycinergic Interneurons_V0 Glutamatergic Interneurons" = "dI6 Glycinergic Interneurons",
                "V2b/s GABAergic/Glycinergic Interneurons_V0 Glutamatergic Interneurons" = "V2b/s GABAergic/Glycinergic Interneurons",
                "KA' GABAergic/Dopaminergic Interneurons_V0 Glutamatergic Interneurons" = "KA' GABAergic/Dopaminergic Interneurons",
                "Intermediate KA Interneurons_V0 Glutamatergic Interneurons" = "Intermediate KA Interneurons",
                "unknown Glycinergic Interneurons_V0 Glutamatergic Interneurons" = "unknown Glycinergic Interneurons",
                "V2a/b Glutamatergic/Cholinergic Interneurons_V0 Glutamatergic Interneurons" = "V2a/b Glutamatergic/Cholinergic Interneurons",
                "Differentiating Interneurons_V0 Glutamatergic Interneurons" = "Differentiating Interneurons",
                "Differentiating Interneurons_Regenerating V0 Glutamatergic Interneurons" = "Differentiating Interneurons",
                "V2a/b Glutamatergic/Cholinergic Interneurons_Regenerating V0 Glutamatergic Interneurons" = "Regenerating V0 Glutamatergic Interneurons")

DimPlot(INs, group.by = "INs_names", label = FALSE)
DimPlot(INs, group.by = "Sample", label = FALSE)

#saveRDS(INs, file = "INs.rds")

INs@meta.data$final_names =
  dplyr::recode(INs@meta.data$info,
                "KA'' GABAergic Interneurons" = "KA'' GABAergic Interneurons",
                "V1 GABAergic/Glycinergic Interneurons" = "V1 GABAergic/Glycinergic Interneurons",
                "V0 Glutamatergic Interneurons" = "V0 Glutamatergic Interneurons",
                "Regenerating V0 Glutamatergic Interneurons" = "V0 Glutamatergic Interneurons",
                "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons" = "V2a/b Glutamatergic/Cholinergic Interneurons",
                "dI6 Glycinergic Interneurons" = "dI6 Glycinergic Interneurons",
                "V2b/s GABAergic/Glycinergic Interneurons" = "V2b/s GABAergic/Glycinergic Interneurons",
                "KA' GABAergic/Dopaminergic Interneurons" = "KA' GABAergic/Dopaminergic Interneurons",
                "Intermediate KA Interneurons" = "Intermediate KA Interneurons",
                "unknown Glycinergic Interneurons" = "unknown Glycinergic Interneurons",
                "V2a/b Glutamatergic/Cholinergic Interneurons" = "V2a/b Glutamatergic/Cholinergic Interneurons",
                "Regenerating V2a/b Glutamatergic/Cholinergic Interneurons" = "V2a/b Glutamatergic/Cholinergic Interneurons",
                "Differentiating Interneurons" = "Differentiating Interneurons")

#saveRDS(neural, file = "neural_final.rds")
```