---
title: "Seurat_neural_annotations_final"
output: html_document
date: "2025-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(ggplot2)
library(dplyr)
library(Seurat)
library(scCustomize)

# Make it so my computer can actually run this code
# options(future.globals.maxSize = 1024^4)
```

Load tail file
```{r}
tail <- LoadSeuratRds(file = "tail_final.rds")
```

First let's do some basic processing on the neural dataset and see how it looks
```{r}
# visualize all tail cell types
DimPlot(tail, reduction = "umap.rpca", group.by = "seurat_cell_types", label = TRUE)

# Subset spinal cord
neural <- subset(x = tail, subset = (seurat_cell_types == "Neural Progenitor Cells" | seurat_cell_types == "Differentiating Neurons" | seurat_cell_types == "Neurons"))

# I could also manually select
#barcodes <- xSelectCells::xSelectCells(tail)
#neural2 <- subset(tail, !(tail@meta.data$orig.ident %in% barcodes))
#neural <- neural2

# Visualize to confirm subset ran
DimPlot(neural, group.by = "seurat_cell_types", label = TRUE)

# Normalize, etc REGRESS OUT SAMPLE AND NFEAT_RNA
neural_scale <- NormalizeData(neural)
neural_scale <- ScaleData(neural_scale, vars.to.regress = c("Sample", "nFeature_RNA"))
neural_scale <- FindVariableFeatures(neural_scale, selection.method = "vst", nfeatures = 2000)
neural_scale <- RunPCA(neural_scale, features = VariableFeatures(object = neural_scale))
neural_scale <- FindNeighbors(neural_scale, dims=1:40, reduction = "pca")
neural_scale <- FindClusters(neural_scale, resolution = 2)
neural_scale <- RunUMAP(neural_scale, dims = 1:40)

# Look at basic paramaters
DimPlot(neural_scale, reduction = "umap", group.by = "seurat_cell_types")
DimPlot(neural_scale, reduction = "umap", group.by = "Sample")
DimPlot(neural_scale, reduction = "umap", group.by = "doublet_finder")

# Definitely needs to be integrated

# Save as you go
#saveRDS(neural, file = "../neural_final.rds")
```

There does seem to be some separation based on the Samples, so I double checked if SCTransform would work better. SCTransform was NOT the vibe- it accentuates the differences between nCount/nFeature/Percent.RP grouping
## DO NOT RUN, BUT INCLUDED FOR CONTEXT ##
```{r}
# SCTransform
#neural_sct <- SCTransform(neural, vars.to.regress = "Sample", verbose = FALSE)

#neural_sct <- RunPCA(neural_sct, features = VariableFeatures(object = neural_sct))
#neural_sct <- RunUMAP(neural_sct, dims = 1:50)

#neural_sct <- FindNeighbors(neural_sct, dims=1:50)
#neural_sct <- FindClusters(neural_sct, resolution = 2)

# Look at basic paramaters
#DimPlot(neural_sct, reduction = "umap", group.by = "cell_types")
#DimPlot(neural_sct, reduction = "umap", group.by = "Sample")
#DimPlot(neural_sct, reduction = "umap", group.by = "doublet_finder")

#FeaturePlot_scCustom(neural_sct, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.rp")
#FeaturePlot_scCustom(neural_sct, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.mt")
#FeaturePlot_scCustom(neural_sct, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nCount_RNA")
#FeaturePlot_scCustom(neural_sct, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nFeature_RNA")
```

Integrate by Sample
```{r}
neural_int <- neural_scale
neural_int[["RNA"]] <- split(neural_int[["RNA"]], f = neural_int$Sample)

neural_int <- IntegrateLayers(object = neural_int, method = RPCAIntegration, orig.reduction = "pca", new.reduction = "integrated.rpca",
    verbose = FALSE)

# re-join layers after integration
neural_int[["RNA"]] <- JoinLayers(neural_int[["RNA"]])

neural_int <- FindNeighbors(neural_int, reduction = "integrated.rpca", dims = 1:40)
neural_int <- FindClusters(neural_int, resolution = 1)
neural_int <- RunUMAP(neural_int, dims = 1:40, reduction = "integrated.rpca")
```

Look at basic parameters
```{r}
DimPlot(neural_int, reduction = "umap", group.by = "seurat_cell_types")
DimPlot(neural_int, reduction = "umap", group.by = "Sample")
DimPlot(neural_int, reduction = "umap", group.by = "doublet_finder")
DimPlot(neural_int, reduction = "umap", group.by = "Phase")

FeaturePlot_scCustom(neural_int, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.rp")
FeaturePlot_scCustom(neural_int, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.mt")
FeaturePlot_scCustom(neural_int, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nCount_RNA")
FeaturePlot_scCustom(neural_int, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nFeature_RNA")

# relabel
neural <- neural_int
neural@meta.data$orig.ident <- rownames(neural@meta.data)

# looking for DRG - nothing in the neural or tail that is clearly neural crest derived DRG
FeaturePlot(neural, features = c("six1", "sox10", "six4", "isl1", "isl2", "sox9", "twist1", "foxd3", "egr2", "myc", "pax3", "tbx2", "snai1", "tbxt", "bmp4", "chrd"))
FeaturePlot(neural, features = c("twist1", "snai1", "bmp4", "pax3", "msx1", "msx2"))
FeaturePlot(tail, features = c("twist1", "snai1", "bmp4", "pax3", "msx1", "msx2"))
FeaturePlot(tail, reduction = "umap.rpca", features = c("scn9a", "scn8a", "scn11a", "prdm12", "trpv1", "trpa1", "piezo2", "foxp2", "ENSXETG00000008680"))
```

```{r}
DimPlot(neural, group.by = "seurat_clusters", label = TRUE)

# take out 0dpa only cluster- probably some very specific type of cell that we can't readily analyze, because it's only in that sample
neural@meta.data$orig.ident <- rownames(neural@meta.data)
barcodes <- xSelectCells::xSelectCells(neural)
neural2 <- subset(neural, !(neural@meta.data$orig.ident %in% barcodes))
neural <- neural2

# take out sclerotome and muscle marked cells - these are probably neuromesodermal progenitors, but likely aren't other things like dorsal root ganglia because they don't express any neural crest markers. These are likely the cells that project with some of the other groups, but were subset with the neural cells because of their cluster identity
markers <- FindAllMarkers(neural, only.pos = TRUE, group.by = "seurat_clusters")
markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
FeaturePlot(neural, features = (top10[which(top10$cluster == "19"),] )$gene)
FeaturePlot(neural, features = (top10[which(top10$cluster == "20"),] )$gene)
FeaturePlot(neural, features = (top10[which(top10$cluster == "22"),] )$gene)

# 19 (tailbud), 20 (some kind of mesenchyme), 22 is muscle
neural2 <- subset(neural, neural@meta.data$seurat_clusters != 19) # tnn, tbxt, lrat, nradd, 
neural2 <- subset(neural, neural@meta.data$seurat_clusters != 20) #pax1, ptx3, pax9, lama4, nid1
neural2 <- subset(neural_sub, neural@meta.data$seurat_clusters != 22) #xirp2, myom2, myot

# take out low quality NPCs, based on percent rp, mt, and RNA counts
FeaturePlot_scCustom(neural_sub, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.rp")
FeaturePlot_scCustom(neural_sub, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "percent.mt")
FeaturePlot_scCustom(neural_sub, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nCount_RNA")
FeaturePlot_scCustom(neural_sub, reduction = "umap", pt.size = 0.5, colors_use = viridis_light_high, features = "nFeature_RNA")
neural_sub <- subset(neural_sub, neural@meta.data$seurat_clusters != 1)

# relabel
neural <- neural_sub
```
Okay I'm actually really happy with that- so the ScaleData and Integration is the way to go! Very clear clusters that I feel represent what I expect to see. Plus, while there's still some residual separation based on nFeature, etc it's much better

Housekeeping and factor samples and save
```{r}
neural@meta.data$orig.ident <- rownames(neural@meta.data)
neural@meta.data$Sample <- factor(neural@meta.data$Sample, levels = c("un0dpa", "un1dpa", "un3dpa", "un7dpa", "0dpa", "1dpa", "3dpa", "7dpa"))

saveRDS(neural, file = "neural_final.rds")
```